The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-20 paper.tex
Work on introduction

\section{Introduction and overview}
\label{sec:intro}

@@

The remainder of this article is organized as follows. 
First, in Section \ref{sec:overview-studies}, we give a brief overview on the various studies that we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2) the Forum study, (3) the Ranger domino study. As an outcome of these studies, we propose two main research hypotheses related to anthropomorphism.
Next, in Section \ref{sec:anthropomorphism} we investigate these hypotheses by synthesizing literature on anthropomorphism. This should also serve to establish a common ground. We first explore an artifact-centered perspective and briefly outline how the human-like design of agents can encourage anthropomorphism. Then, we explore a human-centered perspective by integrating theories from cognitive science, developmental and social psychology that serve to explain anthropomorphism. At the end of this section we bring both perspectives together and interpret that anthropomorphism is a multi-factor phenomenon.
Based on the two previous sections, we propose an initial model of the \textit{``Dynamics of Anthropomorphism''} in Section \ref{sec:dynamics-model}. This model describes the variances and dynamics of people's tendency to anthropomorphize an artificial agent, such as a robot.
Next, in Section \ref{sec:cognition-neuroscience} we present a cognitive and neuroscience perspective on anthropomorphism to consolidate our considerations and the proposed model.
This will be followed by a discussion about the limitations of our approach (Section \ref{sec:discussion}). Further, we debate existing and alternative ways to measure
anthropomorphism in human-robot interactions.
Finally, Section \ref{sec:conclusion} concludes this article and offers starting points for further research on anthropomorphism and the effect of the human-like design approach in robotics.

%{Intro when start with anthropomorphism}
%The essence of anthropomorphism is the perception of human-like characteristics
%in either real or imagined non-human agents. In Human-Robot Interaction (HRI),
%anthropomorphism is a commonly observed phenomenon: people tend to ascribe
%human-like characteristics to robots, and they are willing to treat and react to
%robots as if they were human-social counterparts. The HRI and social robotics
%literature is full of various kinds of studies showing how humans readily
%interact in a social fashion with robots -- despite the fact that robots are not
%humans. In our own studies on HRI analysis in real-world settings, we also
%observed anthropomorphism. However, we found qualitative differences and
%dynamics over time in how participants anthropomorphized and related to the
%robots used in our studies. More concretely, we found that anthropomorphism was
%not only based on how much the robot resembles a human (anthropomorphic design)
%but that it was determined by multiple factors, including individual
%characteristics of the human user but also the context of use, for instance. We also
%found that anthropomorphism changed over time with growing experience of the
%user with the system. By trying to understand these variances and dynamics in
%anthropomorphism, we reviewed related work from the HRI domain and social
%robotics. Despite the fact that there is a vast amount of literature on
%anthropomorphism in HRI, we were surprised to find that other studies do not
%describe similar variances or dynamics. Instead, we came across diverse and
%partially conflicting understandings of anthropomorphism. We also had the
%impression that the reported data were sometimes over-interpreted (probably due
%to the confusion of what exactly anthropomorphism is) or obtained in settings
%and assessed in ways that we believe do not give enough consideration to the
%complex nature of anthropomorphism. However, when reviewing literature from
%social psychology, cognitive science, and neuroscience, we found support for the
%variances and dynamics in anthropomorphism that we observed in our studies. As
%HRI integrates many disciplines, we think that it is valid to use theories from
%these related domains for interpreting our results and that it is relevant to
%report on this integration.
%
%In this article, we would like to share some critical but constructive
%considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
%to a better understanding of this complex phenomenon related to robots. In
%Section \ref{sec:anthropomorphism} we would first like to give an overview of
%how anthropomorphism can be defined, and then, in Section \ref{sec:explanations}
%present explanations to the question why people anthropomorphize, with a focus
%on anthropomorphism as a social phenomenon in HRI and the trend to design
%human-like robots. Then, based on these reviews of theories and concepts from
%related domains, in Section \ref{sec:multi-factors} we propose that
%anthropomorphism is determined and affected by multiple factors. We outline how
%anthropomorphism as a multi-factor phenomenon can impact the user experience and
%acceptance of robots in daily life environments. Next, in Section
%\ref{sec:dynamics-model} we suggest that anthropomorphism changes over time, and
%describe these changes in an initial model that we call ``Dynamics of
%Anthropomorphism''. We illustrate our idea with empirical findings from several
%HRI studies that we conducted in natural environments. In Section
%\ref{sec:cognition-neuroscience} we present a cognitive and neuroscience
%perspective on anthropomorphism to consolidate our considerations and the
%observed dynamics. Next, Section \ref{sec:discussion} considers limitations in
%our approach and we discuss existing and alternative ways to measure
%anthropomorphism in HRI. Finally, in Section \ref{sec:conclusion} we conclude
%this article.

@@

\section{A few results from the field}
\label{sec:overview-studies}

In this section we give a brief overview of the studies that we conducted to investigate anthropomorphism. Note that it was not the main purpose of these studies to analyze in detail the effect of different dimensions human-likeness of a robot on the user's perception of and interaction with the robot. Each of these studies

The strong points

@@

@@

\caption{User perception of the Roomba robot over a 6 month period
($\sigma=15$).}

\label{roomba-perception}
\end{figure*}

\subsection{Forum study: using anthropomorphic language to describe a robot}

\hl{Julia: I think we should also include this study}

@@

@@

\subsection{Towards an understanding of anthropomorphism}

>>>>

\section{Introduction}
\label{sec:intro}

@@

\subsection{A few results from the field}

To support the reflection, it may be useful to consider a few experimental
results. Figure~\ref{roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} robots
were given to nine household for a 6 months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rated
their perception of different features of the robot on a 7-Likert scale. As
Figure~\ref{roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and other decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one do not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed and studied trait of human-robot
interaction, it appears it lacks a formal ground and unified understanding to
take into account the wide range of phenomena that it encompasses.

\subsection{Towards a formal understanding of anthropomorphism}

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Anthropomorphism is traditionally
perceived in robotics as a static feature that once observed during a
short-term interaction reflects a sustaining social effect.  Based on an
literature review previously published~\cite{fink_anthropomorphism_2012}, a
long-term field study in a natural environment~\cite{fink_living_2013}, as well
as two on-going child-robot experiments, we believe that anthropomorphic
effects do evolve over time in non-monotonic ways, and play a central role in
sustaining engagement in human-robot interaction.

We propose to formalize these dynamic effects into a model that we call the
\emph{dynamics of anthropomorphism}. We hope that it may support further
discussions and reflections on how anthropomorphism impacts human-robot
interaction on the long run, and also foster research on the affective bonds
induced by anthropomorphic projections on robots.

As such, we propose to represent how the level of anthropomorphic effects (\ie
observable manifestations of anthropomorphism) evolves over a long-term
human-robot interaction (Figure~\ref{fig:dynamics}). By long-term interaction,
we mean direct (non-mediated), repeated interaction with the same robot, over
an extended period of time (typically longer than a week).

In this article, we would like to share some critical but constructive
considerations on anthropomorphism in robotics and HRI. Thereby, we contribute
to a better understanding of this complex phenomenon related to robots.

The remainder of this article is organized as follows.  First, in Section
\ref{sec:overview-studies}, we give a brief overview on the various studies that
we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2)
the Forum study, (3) the Ranger domino study. As an outcome of these studies, we
propose two main research hypotheses related to anthropomorphism.  Next, in
Section \ref{sec:anthropomorphism} we investigate these hypotheses by
synthesizing literature on anthropomorphism. This should also serve to establish
a common ground. We first explore an artifact-centered perspective and briefly
outline how the human-like design of agents can encourage anthropomorphism.
Then, we explore a human-centered perspective by integrating theories from
cognitive science, developmental and social psychology that serve to explain
anthropomorphism. At the end of this section we bring both perspectives together
and interpret that anthropomorphism is a multi-factor phenomenon.  Based on the
two previous sections, we propose an initial model of the \textit{``Dynamics of
Anthropomorphism''} in Section \ref{sec:our-ideas}. This model describes the
variances and dynamics of people's tendency to anthropomorphize an artificial
agent, such as a robot.  Next, in Section \ref{sec:cognition-neuroscience} we
present a cognitive and neuroscience perspective on anthropomorphism to
consolidate our considerations and the proposed model.  This will be followed by
a discussion about the limitations of our approach (Section
\ref{sec:discussion}). Further, we debate existing and alternative ways to
measure anthropomorphism in human-robot interactions.  Finally, Section
\ref{sec:conclusion} concludes this article and offers starting points for
further research on anthropomorphism and the effect of the human-like design
approach in robotics.

@@

\section{Studying anthropomorphism}
\label{sec:overview-studies}

In this section we give a brief overview of the studies that we conducted to
investigate anthropomorphism in human-robot interaction. Note that it was not
the main purpose of these studies to analyze in detail the effect of different
dimensions human-likeness of a robot on the user's perception of and interaction
with the robot. Each of these studies were motivated also by other research
questions, and we encourage the reader to refer to our other main publications
about these studies for more details.

@@

We openly explored long-term interactions between human users and a domestic
service robot in a 6-months lasting ethnographic study. We gave a commercially
available vacuum cleaning robot (iRobot's Roomba, see Figure
\ref{fig:aibo-roomba-ipad} middle) to nine households in the French speaking
part of Switzerland. The fairly diverse sample of households consisted of 2
single homes, 1 couple, and 6 families with children in their home. In total,
this summed up to 30 participants: 15 adults (26-71 years old, mean age 43.6
years), and 15 children (6 months to 18 years old). During the 6 months, each
household was visited 5 times to conduct interviews and make field observations
in the households. Moreover, during several weeks, households were asked to fill
in a diary to report the usage of the robot and other cleaning activities.
\footnote{This study is published in \cite{fink_living_2013}.} Hereafter, we
refer to this study as the "Roomba study".

%The main goal of this study was to explore how households live with a vacuum
%cleaning robot over an extended period of time. We wanted to analyze how the
%usage of and interaction with the robot evolve over time, and to identify
%adoption phases as well as factors that impact on the long-term acceptance of
%service robots in homes.

From the Roomba study, we can report on several findings related to
anthropomorphism and the anthropomorphic design of robots.
%The gathered data for these aspects include qualitative observations,
%semi-structured interviews, and also quantitative assessments of various
%aspects of human-robot interaction.
The Roomba study is one of very few long-term human-robot interaction studies in
the domestic environment.

In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.\footnote{\hl{note that these two different kinds of data
(behavior/interaction and perception) already show that we try to measure
anthropomorphism from these two angles!}}

\subsubsection{Anthropomorphic interactions with the robot\\}

It has been described in several previous studies of the Roomba that users tend
to engage in a variety of non-cleaning activities associated with the robot,
including, but not limited to, naming, ascribing personality and gender
\citep{forlizzi_service_2006,forlizzi_how_2007,krumm_my_2007,sung_domestic_2010}.
\cite{sung_housewives_2008} carried out a survey among 379 Roomba users to
investigate how common these activities are. The authors concluded that overall,
respondents had relatively low rates of treating Roomba as a living entity.

In our long-term study of the Roomba in homes, we found that it was mostly
during the initial phase with the robot that some (rather few) people tended to
interact in an anthropomorphic or human-social way with their vacuum cleaning
robot.  Overall, our data suggest that the anthropomorphic effect and the social
impact of a functional robot on a household is rather marginal.
%There were just few instances of people anthropomorphizing the robot and
%interacting with it in a human-social way. More importantly, overall, these
%anthropomorphic interactions with the robot did not last but seemed to be
%somehow related to the novelty of the robot.

We categorized several interactions and actions with the robot as
anthropomorphic or human-social (in a broader sense). 
%Examples of what we categorized as somehow social and anthropomorphic
%(human-like) interactions with the robot include \textit{watching} the robot
%for fun, giving a \textit{demo} of the robot to guests / visitors, taking a
%\textit{photo/video} of the robot, \textit{personalizing} the robot, giving a
%\textit{name} to the robot, \textit{talking} to the robot using direct speech,
%\textit{playing} with the robot, and making \textit{adjustments} to the home
%due to the robot. 
Figure \ref{fig:roomba-activities}) shows how many of the 9 households engaged
in these activities during the 6 months when the study was conducted.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.}

\label{fig:roomba-activities}
\end{figure}

\paragraph*{Watching the robot for fun.} We categorize watching a robot for fun
as an entertaining, emotional activity, reflecting a feeling of empathy for the
robot. In \cite{sung_housewives_2008} survey, 73~\% of the participants reported
to watch their Roomba running for fun. While in the beginning of our study, also
a quite high proportion (6 of 9 households) reported to watch the robot for fun,
after 2 and 6 months the number of households that engaged in this activity
decreased by half.

\paragraph*{Giving a robot demo to visitors.} It is somehow natural that owners
of innovative devices show the device to others and ask for their opinion about
it \citep{rogers_diffusion_1995}. However, this activity suggests that the robot
is presented as a new part of the household to visitors.
\cite{sung_domestic_2010} described that a vacuum cleaning robot can serve as a
mediator and encourage social interaction among people. We categorize giving a
demo of the robot to visitors / guests as such an activity. During the first 2
weeks with the robot 8 of 9 households in our study showed their robot to
others, and still after 2 and 6 months 5 of the households engaged in this
activity. This number is more or less in accordance with the 57~\% of survey
respondents who stated to give demonstrations to others
\citep{sung_housewives_2008}.

%\paragraph*{Taking photos / videos of the robot.} How far taking a photo or
%video of a robot can be seen as an anthropomorphic interaction is questionable.
%However, it reflects the fact that the robot is observed by the user who for
%some reason is motivated to record the event. The number of households that
%took a photo or video of the robot decreased over time from 5 households in the
%beginning to 1 household after 6 months.

\paragraph*{Personalizing the robot.} It has been found that people are
motivated to personalize technology as a form of self-expression (identity
projection) or when they use a technology a lot and know how to customize it
(technical identity projection) \citep{blom_personalization:_2000}. Though
\cite{sung_pimp_2009} found that some Roomba users who were given a
customization toolkit personalized their Roomba, none of the users in her study
who was \textit{not} given this toolkit personalized their Roomba. Also in our
study none of the 9 households engaged in a personalization activity.

\paragraph*{Giving a name to the robot.} Giving a name to an artificial agent
reflects a certain kind of personal relationship to the agent. By naming a
robot, it is not anonymous anymore, and might be perceived as a moral agent. As
such, naming an artificial agent can be classified as anthropomorphizing this
agent. It further suggests that the agent is talked to directly. Though more
than 23~\% of the survey participants in \cite{sung_housewives_2008} study
reported to name their Roomba, only 1 of 9 households in our study did so.

\paragraph*{Talking directly to the robot.} Using direct speech toward an object
or artificial agent that does not have the ability to recognize and process
speech (such as verbal comments) or other sound cues, is an interesting
phenomenon which can be classified as an anthropomorphic interaction. Whereas
only 11~\% of the survey respondents in \cite{sung_housewives_2008} study
reported to talk to their robot, around half of the households engaged in this
activity and the proportion did not notably decrease over time.

\paragraph*{Playing with the robot.} In our study, it were mostly children who
engaged in playing with the robot, however, most of them stopped doing so after
some time. The number of household playing with their Roomba decreased from 8
when asked after 2 weeks with the robot, to 4 households after 2 months, and 1
household after 6 months. In \cite{sung_housewives_2008} survey, 37~\% of the
Roomba owners reported to play and experiment with their robot.\\

%\paragraph*{Making adjustments to the home due to the robot.} To help Roomba
%clean, \cite{sung_housewives_2008} found in an online survey that about half of
%the 379 participants made physical modifications to their homes. This process
%is referred to as \textit{roombarization}, and it seems to be a unique aspect
%of adopting a domestic service robot. In our study, almost all the households
%made adjustments when they used the robot for the first time in front of us.
%However, during the first two weeks with the robot, only 3 households reported
%to have adjusted something due to the robot, and after 6 months it was only 1
%household that still engaged in this activity. This decrease is, however,
%logical since after some time most people had done all the necessary
%adjustments to enable the robot show a good performance.

%\paragraph*{Emerging anthropomorphic interactions.} Another, from our side more
%unexpected interaction that we would describe as somehow anthropomorphic, is
%what we call \textit{``collaborative cleaning''} with the robot. In the initial
%phase of the study, some people spontaneously helped the robot to clean by
%either picking up crumbs from the floor themselves or pushing it right in front
%of the robot. Another way of collaborating with the robot we observed was to
%help it go everywhere by spontaneously putting up the trash bin or the carpet
%in the bathroom (things that are not just lying around and need to be cleaned
%up). However, the activity of collaborative cleaning faded out soon, and we
%suspect it was encouraged by the novelty of the robot.\\

Overall, Figure \ref{fig:roomba-activities} shows that as time goes by, less
households engage in the before-mentioned non-cleaning activities with the
vacuum cleaning robot: after 2 weeks, on average, 4.67 of 9 households engage in
a human-like activity with the robot; after 2 months, the value decreases to
2.83, and after 6 months to 2.3. Our interpretation is that the nature of
human-robot interaction changes over time. More concretely, we suspect  that
anthropomorphic / human-social interactions with the robot become less over
time, and that only with some people these interactions sustain.

\subsubsection{Perception of anthropomorphic characteristics of the robot\\}

We also studied people's perception of the robot, and how it evolved over time
(see Figure \ref{fig:roomba-perception}). The main idea was to find out which of
the perceived attributes could be related to how far the robot was integrated in
the household as a cleaning tool. Concerning anthropomorphism, we included
\textit{perceived intelligence} and experienced \textit{fun} when using the
robot as indicators for the perception of human-likeness in the robot.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

@@

    \caption{People's perception of the robot over time, n = 15 participants.
    This line chart shows how people's perception of the robot evolves over
    time. At 5 time points people were asked to rate the robot according to the
    following aspects: the perceived \texttt{intelligence} of the robot, the
    perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much
    \texttt{fun} it was to use the robot, how far they found the robot had an
    \texttt{impact} on their household, and their {overall} impression of it.}

    \label{fig:roomba-perception}
\end{figure}

\paragraph*{Perceived intelligence.} On a scale from 1-7, people were not sure
about how intelligent they should expect the robot to be (mean value 3.9). The
rating did not change much and not significantly over time (average ratings
between 3.3-4.2) \citep{fink_living_2013}. Interestingly, at all five time
points, male participants rated the robot as more intelligent (average 4.5) than
females (average 3.4), however, this difference was not significant. When asked
why they thought Roomba possesses intelligence, men more often referred to the
'robot side' of Roomba: its abilities to detect and avoid obstacles, stairs, and
being able to recharge itself. Contrary, women referred to its rather limited
vacuuming performance and found it was not as intelligent because \textit{``it
doesn't see where the dirt is, it sometimes leaves it, it can't be smart!''}.
This difference in the perception of intelligence of a robot suggests that
gender,and probably other individual characteristics need to be taken into
account.

\paragraph*{Experienced fun.} Participants rated the robot as 'somewhat fun'
(values from 4.0-4.6). There was a qualitative gender difference (not
significant), in the sense that the experienced fun remained quite stable for
male participants, whereas it tended to decrease with time for female
participants. Men described the robot more as a 'fun gadget' than females, who
referred to it as \textit{``In the end, Roomba is a cleaning tool and it is not
really fun to use it''}. Again for this aspect, gender seems to play a role in
how the robot is perceived and experienced.\\

Overall, Figure \ref{fig:roomba-perception} shows that after the first usage of
the Roomba (T2), most of the perceived constructs slightly decrease (and hence
are less positively rated) as time goes by. The course is different for the
perceived \textit{ease of use} of the robot, as participants become more
experienced with the robot over time. Also, the experienced \textit{fun} does
not decrease with time.

\subsubsection{Summary Roomba study\\}

In accordance with previous work we also found relatively low rates of users
treating and perceiving their vacuum cleaning robot as a living (or human-like)
entity \citep{sung_housewives_2008}. We interpret that it might be the pure
functional purpose and non human-like form of the Roomba that prevented people
from anthropomorphizing it. Still, we observed some initial anthropomorphic
effects, both reflected in interaction with and perception of the robot.
However, most of the anthropomorphic effects did not last long. \hl{One fairly
stable human-like interaction with the robot was \textit{talking to the robot},
despite the fact that the robot was not able to recognize and respond to verbal
statements.\footnote{important! I think I've seen that finding somewhere else
when comparing ASIMO to AIBO or so.}}

\subsection{Forum study: using anthropomorphic expressions to describe a robot}

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

We investigated the amount and context of anthropomorphic language in three
online discussion forums about the robotic dog AIBO, the vacuum cleaning robot
Roomba, and the tablet computer iPad. This was done by means of a content
analysis of 750 forum posts that were further partitioned in n = 1208 valid
segments \citep{fink_anthropomorphic_2012}. Hereafter, we refer to this study as
the "Forum study". The three devices, pictured in Figure
\ref{fig:aibo-roomba-ipad}, were chosen as they show different designs and
fulfill different purposes. We would describe (1) AIBO as a zoomorphic /
pet-like robot that does not have a clear purpose but engages people to play
with it; (2) Roomba as a mechanical / functional domestic service robot that has
a practical main purpose (vacuum cleaning) but offers not much of interactivity;
(3) the iPad as a minimalistic designed interactive (touch-based) technology
with multiple purposes (both entertainment, and professional / practical use).
Based on these different characteristics, we made the two following hypotheses
concerning anthropomorphism:

    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.

    	\item As anthropomorphism can be seen as a certain kind of emotional / social relation to an artifact, there will be more anthropomorphic language in forum posts in which the author describes a relation to the device, compared to discussing technological aspects or how the device is used.
\end{enumerate}

Similar to how \cite{friedman_hardware_2003} analyzed anthropomorphic language
in online forums, we coded a segment as anthropomorphic when the respective
device was described in terms of:

\begin{itemize}
    \item \textit{Life-likeness}, uch as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};

    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};

    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};

    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};

    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};

    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.

\end{itemize}

In the next sections, we present findings in respect of our two hypotheses:
anthropomorphic language related to the device, and anthropomorphic language
related to the topic of the forum post.

Overall, our analysis revealed that in all three forums, discussions about
\textit{technical} aspects of the respective device were most frequent (67.7~\%
of all 1208 categorized segments). However there were huge differences between
the three devices: \textit{technical} aspects made up 79.9~\% of the discussion
in the Roomba forum, compared to 69.9~\% in the AIBO forum, and 55.2~\% in the
iPad forum. Another fairly frequently discussed topic concerned the
\textit{usage} of the respective device, thus people describing how and for what
they use their technology. This topic made up 21.8~\% of all posts, and was most
frequent in the iPad forum (35.7~\%), followed by the Roomba forum (15.9~\%),
and the AIBO forum (10.7~\%). Last but not least, we defined \textit{relation}
as the third topic of interest. Discussions in which people described their
relation to the respective device made up 10.5~\% of all posts, and was most
frequent in the AIBO forum (19.4~\%), followed by the iPad forum (9.1~\%), and
the Roomba forum (4.2~\%). \hl{(if it is getting too much we can leave out this
paragraph, it's not crucial, and doesn't say anything about anthropomorphism
directly)}

\subsubsection{Anthropomorphic language related to the device (design)\\}

A statistical analysis revealed a significant difference for the use of
anthropomorphism across the three devices \citep{fink_anthropomorphic_2012}. In
accordance with our first hypothesis, the highest amount of anthropomorphic
language (56.6~\%) was found in the AIBO forum. In contrast, only 11.7~\% and
respectively 2.2~\% of the segments in the Roomba and the iPad forum contained
anthropomorphic expressions.  The data suggests that on a subjective scale of
perceived human-likeness (or rather life-likeness), Roomba is closer to the iPad
than to the AIBO, at least in matters of how they are described with words. This
was a bit against our expectation. Still, the Roomba is an autonomous robot that
can navigate around the house itself, and we expected this property to enhance
anthropomorphism. We interpret that here, the physical embodiment of the object
and how far it resembles a life-like entity determine how far the object is
anthropomorphized, and as such the robotic dog is anthropomorphized more than
the functionally designed devices.

\subsubsection{Anthropomorphic language related to the topic of conversation\\}

Also in terms of anthropomorphic language related to the topic of conversation,
a statistical analysis revealed a significant difference
\citep{fink_anthropomorphic_2012}. The analysis suggests that anthropomorphizing
happens more when the topic of conversation is \textit{relation}, compared to
\textit{usage}, and \textit{technical} aspects. This finding is qualitatively
confirmed when looking at the distribution of anthropomorphic expressions across
the three topics in the Roomba and iPad forum (with both technologies the
highest amount of anthropomorphism is in the \textit{relation} topic). However,
in the AIBO the amount of anthropomorphic segments is highest for the
\textit{usage} topic.  We interpret that it is the different type of interaction
/ use context that makes the difference. Contrary to the iPad and the Roomba,
the AIBO is made to encourage its user in a playful / social interaction, which
in turn is a typical use context. Furhter, AIBO has the ability to respond
intelligently to its user and environment, which might also encourage people to
ascribe human-like abilities to it while interacting with it.

\subsubsection{Summary Forum study\\}

Overall, findings of the content analysis suggest that there are differences in
people's tendency to use anthropomorphic language, originating form the design
characteristics of the artifact. Interestingly, not only the physical embodiment
of the object but also the interaction complexity and purpose of use of the
object seems to play a role in how far the object is anthropomorphized.

@@

The so-called "Ranger domino study" describes a child-robot interaction study
that we carried out in a fairly controlled lab setting. One of the goals of this
study was to investigate the effect of unexpected robot behavior on (1) the
interaction and engagement with robot, and (2) how far children would ascribe
human-like abilities (\eg intentions, cognitive abilities) to the robot.

13 groups of each two children (n = 26) who were all between 4-5 years old and
knew each other, were invited to play for around 20 minutes with the Ranger
robot toy box \citep{fink2014which} and assemble a domino together. The robot
was operated by one of the experimenters using the wizard-of-oz technique.
Ranger was supposed to transport domino tiles between the two children and it
would do its job correctly for the first 7 domino tiles. Then, however, the
robot started misbehaving by either \textit{disobeying} to the child

@@
